---
title: "Monthly_Average_Salary_Analysis"
author: "Wharton Chan"
date: "17 February 2016"
output: pdf_document
geometry: margin = 0.5in
---
```{r, echo = FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60))
```

\section{Pre-Processing}
First load required libraries:  
```{r libraries, warning = FALSE,message=FALSE}
library(ggplot2)
library(RColorBrewer)
library(data.table)
library(reshape2)
```

Load in data, which has been pre-processed in Excel to give a .csv  
```{r}
setwd("~/Gary/Data_Analysis")
data_wage = read.csv("monthly_average_salary.csv",header = T, as.is = T,na.strings = "N.A.")
```

Create column for time objects for time series plotting:  
```{r}
date_cols = data_wage[,1:2]
date_cols$conc_date = paste(date_cols$Year,date_cols$Month,"01",sep = " ")
date_cols$date = as.Date(date_cols$conc_date,"%Y %b %d")
data_wage = cbind(date_cols$date,data_wage[,3:length(data_wage)])
colnames(data_wage)[1] = "Date"
```

From the raw data the occupations are also subdivided into groups. Create such list:  
```{r occ_list,tidy=TRUE}
occ_list = list()
occ_list[["sup_tech"]] = c("office_supervisor","account_supervisor","estate_officer")
occ_list[["clerical"]] = c("stock_clerk","account_clerk","general_clerk","cashier_clerk","receptionist","shipping_clerk","restaurant_receptionist")
occ_list[["service"]] = c("cook","waiter","security_guard")
occ_list[["misc"]] = c("general_worker","office_assistant","driver","dishwasher","lav_cleaner","gen_cleaner")
```

\subsection{Account for Inflation}
Load in data for CPI, and create time objects similarly:  
```{r}
data_cpi = read.csv("CPI_monthly_1990_2015.csv",,header = T, as.is = T,na.strings = "N.A.")
date_cols = data_cpi[,1:2]
date_cols$conc_date = paste(date_cols$Year,date_cols$Month,"01",sep = " ")
date_cols$date = as.Date(date_cols$conc_date,"%Y %b %d")
data_cpi = cbind(date_cols$date,data_cpi[,3:length(data_cpi)])
colnames(data_cpi)[1] = "Date"
```

To calculate real prices, we align the prices to the latest CPI, which will be the 20150901 timepoint. It has been decided that basket A will be used (temporarily) given the low income nature of the list of occupations within the dataset.  
```{r CPI_A, tidy = TRUE}
curr_cpi = data_cpi[which(data_cpi$Date == "2015-09-01"),"A_index"]
data_cpi$A_ratio = curr_cpi/data_cpi$A_index
data_adj = data_cpi[,c("Date","A_ratio")]
data_wage2 = merge(data_wage,data_adj,by="Date",all.x = T)
#multiply prices with column A_ratio
data_wage2 = data.table(data_wage2)
data_wage_adjA = data_wage2[,.(office_supervisor = office_supervisor*A_ratio,account_supervisor = account_supervisor*A_ratio,estate_officer = estate_officer*A_ratio, stock_clerk = stock_clerk*A_ratio,account_clerk = account_clerk*A_ratio,general_clerk = general_clerk*A_ratio,cashier_clerk = cashier_clerk*A_ratio,receptionist = receptionist*A_ratio,shipping_clerk = shipping_clerk*A_ratio,restaurant_receptionist = restaurant_receptionist*A_ratio,cook=cook*A_ratio,waiter=waiter*A_ratio,security_guard=security_guard*A_ratio,general_worker=general_worker*A_ratio,office_assistant=office_assistant*A_ratio,driver=driver*A_ratio,dishwasher=dishwasher*A_ratio,lav_cleaner=lav_cleaner*A_ratio,gen_cleaner=gen_cleaner*A_ratio),by=Date]
```

To compare the best CPI to use, we also try composite CPI:  
```{r CPI_comp,tidy=TRUE}
curr_cpi = data_cpi[which(data_cpi$Date == "2015-09-01"),"comp_index"]
data_cpi$comp_ratio = curr_cpi/data_cpi$comp_index
data_adj = data_cpi[,c("Date","comp_ratio")]
data_wage2 = merge(data_wage,data_adj,by="Date",all.x = T)
#multiply prices with column comp_ratio
data_wage2 = data.table(data_wage2)
data_wage_adjcomp = data_wage2[,.(office_supervisor = office_supervisor*comp_ratio,account_supervisor = account_supervisor*comp_ratio,estate_officer = estate_officer*comp_ratio, stock_clerk = stock_clerk*comp_ratio,account_clerk = account_clerk*comp_ratio,general_clerk = general_clerk*comp_ratio,cashier_clerk = cashier_clerk*comp_ratio,receptionist = receptionist*comp_ratio,shipping_clerk = shipping_clerk*comp_ratio,restaurant_receptionist = restaurant_receptionist*comp_ratio,cook=cook*comp_ratio,waiter=waiter*comp_ratio,security_guard=security_guard*comp_ratio,general_worker=general_worker*comp_ratio,office_assistant=office_assistant*comp_ratio,driver=driver*comp_ratio,dishwasher=dishwasher*comp_ratio,lav_cleaner=lav_cleaner*comp_ratio,gen_cleaner=gen_cleaner*comp_ratio),by=Date]
```

Now change data structure in both data frames to allow efficient plotting:  
```{r melting}
adjcomp = melt(data_wage_adjcomp,id.var = "Date")
adjA = melt(data_wage_adjA,id.var = "Date")
colnames(adjcomp) = c("Date","Occupation","Salary")
colnames(adjA) = c("Date","Occupation","Salary")
```
\newpage

\section{Plots}

There are 19 occupations, so first define the colours:  
```{r colours}
set_colour = c(brewer.pal(12,"Paired"),brewer.pal(7,"Set3"))
names(set_colour) = colnames(data_wage_adjA)[2:length(data_wage_adjA)]
```

Plot for df adjusted with basket A:  
```{r A_plot,fig.width = 8,warning = FALSE,tidy = TRUE,fig.path='figure/'}
ggplot(adjA, aes(x=Date, y=Salary, colour=Occupation)) + geom_line() + scale_x_date() + labs(x = "Time", y ="Real Average Salary",title = "Average Real Salary in Hong Kong for selected Occupations CPI_A") + scale_colour_manual("variable",values = set_colour) + theme_bw() + theme(legend.position="bottom",plot.title = element_text(size = rel(0.9))) + guides(colour=guide_legend(nrow=4,byrow=TRUE,title=NULL))
```

Plot for df adjusted with composite basket:  
```{r comp_plot,fig.width = 8,warning = FALSE,tidy = TRUE,fig.path='figure/'}
ggplot(adjcomp, aes(x=Date, y=Salary, colour=Occupation)) + geom_line() +
  scale_x_date() + labs(x = "Time", y ="Real Average Salary",title = "Average Real Salary in Hong Kong for selected Occupations CPI_Comp") + scale_colour_manual("variable",values = set_colour) + theme_bw() + theme(legend.position="bottom",plot.title = element_text(size = rel(0.9))) + guides(colour=guide_legend(nrow=4,byrow=TRUE,title=NULL))
```





